Investigation of ant cuticle dataset using image texture analysis methods

### 2022-07-22 paper/main.tex
hung changes

\title{Ant cuticle image classification using texture analysis: a comparative
    study}

@@

There is a large variety of ant species and most species are diverse in terms of
size, shape, behaviors, and cuticle textures \colorbox{green}{[references]}.
Based on our initial observation of the cuticle sculpturing in ants, it appears
that some type of textural structures can be identified in images. Therefore,
this image dataset may be worth to be explored from the image texture analysis
point of view for the further knowledge discovery. Image texture analysis has
been used in image processing for many decades \colorbox{green}{[references]}. Due to its
diversity of images on scales, smoothness, coarseness, microtexture, and
macrotexture occurring in images, it is a difficult problem for image textural
interpreters. With a rapid development of deep machine learning, deep neural
networks have been frequently used in image texture analysis \colorbox{green}{[references]}.

Our contribution is to establish an ant image dataset which can be used to
explore the categorization of different ant cuticle textures using image texture
analysis methods. \colorbox{green}{(comment: we may need to say how the dataset
    is established)} Our experiment shows that the deep learning networks with image
texture embedded in the design give a better accuracy than the K-views
clustering method.

@@

Taxonomists have developed extensive terminology describing ant cuticle
microsculpture as it is often a useful diagnostic trait to distinguish between
closely related ant species \cite{blaimer_taxonomy_2019,fisher_ants_2007}. The
definitive text on ant cuticle terminology — \textit{The Glossary of Surface
    Sculpturing} — contains over 100 terms to describe the cuticle sculpturing
patterns of ants \cite{harris_glossary_1979}. Despite the numerous ways to
describe microsculpture, in general, microsculpture diversity falls into two
classes - smooth and rough. Smooth ants are free of any obvious microsculpture
and may have small hairs that give the cuticle a dull appearance. Rough ants
have a variety of raised lines or depressed textures on the cuticle. The
entirety of ant microsculpture is represented within these two classes.

Cuticle sculpturing in ants has been explored thoroughly from a taxonomic
perspective; however, the function of nano and microstructures on insect
exoskeletons is a developing topic in entomology. Watson et al. review the
literature of cuticle nano and microstructure function and propose 21 possible
functions associated with these structures \cite{watson_diversity_2017}. Many of
these functions are related to structures not found in ants such as scales and
nanostructures. The review does include functions that may relate to ants such
as friction control, enhanced surface area, and increased hardness. Watson et
al. also describe seven types of cuticle structures ranging from hairs and
scales to nano and micro structures \cite{watson_diversity_2017}. The cuticle
sculpturing of ants seems to fall within one type - complex microstructures.

The five broad functional groupings developed that describe the complex
microstructures found on ants were derived from reviewing the variation of
cuticle sculpturing across ants. These functional groupings were also reflected
in Harris as many terms could be grouped together based on similar definitions
and comparing the scanning electron microscope photographs provided in the
publication \cite{harris_glossary_1979}.

There is often substantial overlap among terms, and closely related cuticle
patterns are likely to be functionally similar. For example, the definition for
“imbricate” is, “partly overlapping and appearing like shingles on a roof or
scales on a fish,” which is difficult to distinguish from the definition of
tesselate, “made up of squares like a chess board, either in sculpturing or in
color.” However, we note that the terms used to describe the cuticle patterns
are similar to the terms used to describe textural patterns in texture analysis.

@@

\begin{table}[h]
    \centering
    \caption{Summary of the dataset.}
    \begin{tabular}{lc}
        \toprule
        Total images  & 2499    \\
        Rough images  & 1072    \\
        Smooth images & 1427    \\
        Image size    & 256x256 \\
    \end{tabular}
    \label{tab:dataset-summary}
\end{table}

@@

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
        \caption*{Smooth texture}
        \label{fig:correct_ideal_518}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
        \caption*{Smooth texture}
        \label{fig:correct_ideal_808}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
        \caption*{Rough texture}
        \label{fig:correct_ideal_842}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
        \caption*{Rough texture}
        \label{fig:correct_ideal_1091}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_346}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_1554}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_1694}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
        \caption*{Smooth texture}
        \label{fig:correct_nonideal_388}
    \end{subfigure}
    \caption{Correctly classified images.}
    \label{fig:correct_images}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
        \caption*{Rough texture}
        \label{fig:incorrect_ideal_84}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_86}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_177}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_197}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_1}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_22}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_61}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_204}
    \end{subfigure}
    \caption{Incorrectly classified images.}
    \label{fig:incorrect_images}
\end{figure}

>>>>

\title{Investigation of ant cuticle dataset using image texture
    analysis methods}

@@

There is a large variety of ant species and most species are diverse in terms of
size, shape, behaviors, and cuticle textures \cite{harris_glossary_1979}.
Based on our initial observation of the cuticle sculpturing in ants, it appears
that some type of textural structures can be identified in images. Therefore,
this image dataset may be worth to be explored from the image texture analysis
point of view for the further knowledge discovery. Image texture analysis has
been used in image processing for many decades \cite{hung_image_2019}. Due to its
diversity of images on scales, smoothness, coarseness, microtexture, and
macrotexture occurring in images, it is a difficult problem for image textural
interpreters. With a rapid development of deep machine learning, deep neural
networks have been frequently used in image texture analysis
\cite{sun_robust_2014,liu_bow_2019,liu_pestnet_2019}.

Our contribution is to establish an ant image dataset which can be used to
explore the categorization of different ant cuticle textures using image texture
analysis methods. We share how the dataset of ant cuticle images is established
in Section 3. Our experiment shows that the deep learning networks designed for
image texture analysis give a better accuracy than the K-views clustering
method.

@@

Because ants lack bold coloration and complex wing venation patterns used to
identify other insect species, microsculpturing has taken a prominent role in
ant identification. To aid in species identification, ant taxonomists have
developed nearly 100 terms to describe  fine differences in microsculpturing
patterns \cite{harris_glossary_1979}. Despite the prominent role of
microsculpturing in ant identification, however, almost nothing is known about
how these patterns evolved or their function. Similar microsculpturing patterns
in other organisms have been found to reduce abrasion, increase stiffness, and
perform as structural antibiotics \cite{zhiwu_erosion_2012, han_efficient_2015,
    rees_form_1975, chung_impact_2007,hasan_selective_2013}. Developing a better
understanding of the function of microsculpturing in ants could provide insight
into their evolution and biology as well as lead to the development of
bio-inspired technologies \cite{penick_comparative_2022}.

One challenge for functional investigations of cuticle microsculpturing is the
complexity of terms used to describe these patterns. Complex terminology may be
useful for describing fine differences between closely related species, but
broader comparisons require a simplified classification scheme. We recently
developed a classification system for microsculpturing that consists of only
five categories: smooth, striate, punctate, reticulate, and tuberous. Using
these categories to classify 11,722 ant species, we found that the earliest ants
were likely smooth and that cuticle microsculpturing independently evolved later
in multiple lineages. An even simpler classification scheme is to divide ants
into only two categories: smooth and rough. Because microsculptuing has evolved
numerous times in ants, this binary classification scheme is sufficient to
detect conditions that select for any sculpturing at all compared to a smooth
state.

Using either a simplified or binary classification scheme, the next challenge is
figuring out how to apply either of these schemes to large numbers of species.
We live in an era of big data, and insect researchers have access to millions of
images available through digital museum collections \cite{beaman_mass_2012}.
Going through these digital collections manually is not feasible for large
projects. Focusing only on ants, for example, there are over 14,000 described
species, and each species has multiple worker and reproductive castes that may
vary in their microsculpturing patterns. Even within an individual ant,
microsculpturing typically varies among and within body segments, which
exponentially increases the number of images that must be assessed. Applied
computing offers a solution to this problem through developing tools that
automate trait classification. Automated approaches would allow researchers to
leverage existing image repositories to extract large trait datasets that can
then be used to detect evolutionary patterns.

@@

\begin{table}[h]
    \centering
    \caption{Summary of the dataset.}
    \begin{tabular}{|l|c|}
        \toprule
        Total images  & 2499    \\
        Rough images  & 1072    \\
        Smooth images & 1427    \\
        Image size    & 256x256 \\
        \bottomrule
    \end{tabular}
    \label{tab:dataset-summary}
\end{table}

@@

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
        \caption*{Smooth texture}
        \label{fig:correct_ideal_518}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
        \caption*{Smooth texture}
        \label{fig:correct_ideal_808}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
        \caption*{Rough texture}
        \label{fig:correct_ideal_842}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
        \caption*{Rough texture}
        \label{fig:correct_ideal_1091}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_346}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_1554}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
        \caption*{Rough texture}
        \label{fig:correct_nonideal_1694}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
        \caption*{Smooth texture}
        \label{fig:correct_nonideal_388}
    \end{subfigure}
    \caption{Correctly classified images from fine-tuned ResNet101.}
    \label{fig:correct_images}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
        \caption*{Rough texture}
        \label{fig:incorrect_ideal_84}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_86}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_177}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
        \caption*{Smooth texture}
        \label{fig:incorrect_ideal_197}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_1}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_22}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_61}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
        \caption*{Rough texture}
        \label{fig:incorrect_nonideal_204}
    \end{subfigure}
    \caption{Incorrectly classified images from fine-tuned ResNet101.}
    \label{fig:incorrect_images}
\end{figure}

\newcommand{\segmentedsubwidth}{0.24\textwidth}
\begin{figure}
    \centering
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r103.png}
        \caption*{Rough texture}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r111.png}
        \caption*{Smooth texture}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s105.png}
        \caption*{Rough texture}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s107.png}
        \caption*{Smooth texture}
    \end{subfigure}

    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/sr103.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/sr111.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/ss105.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/ss107.png}
    \end{subfigure}

    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat103.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat111.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat105.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat107.png}
    \end{subfigure}
    \caption{Raw images are shown with corresponding segmented outputs from
        K-views algorithm. The second row shows the results of using size of
        15x15 patches. The third row shows the results of using size of 19x19
        patches.}
    \label{fig:kviews_segmented}
\end{figure}

\FloatBarrier Some classified results from the K-views are shown in Figure
\ref{fig:kviews_segmented}. A training set of 500 patches for background, 500
patches for rough, and 500 patches for smooth are applied with patch sizes of
15x15 and 19x19. These patches are extracted from two rough and two smooth
images. The K-views algorithm converges early although we set the maximum number
of iterations to be 1000.
